A {practical} review on linear and nonlinear global approaches to stability analysis

### 2018-05-24 ARTICLE_STABFEM/ARTICLE_ASME_V2.tex
version 2 of the paper

\begin{figure}
\begin{center}
\includegraphics[width=.9 \linewidth]{Cylinder_SC100_EnergySigma.png}
\includegraphics[width=.9 \linewidth]{Cylinder_SC100_CySigma.png}
\end{center}
\caption{Self-consistent approach for the wake of a cyrcular cilinder: results at $Re = 
100$.}
\label{fig:SC100}
\end{figure}
We now briefly review the self-consistent approach as introduced by Manti\v{c}-Lugo et 
al \cite{MLugo2014}. The authors adopted a pseudo-eigenmode expansion of the flow 
as follows: 
\be{SC}
{\bf u } = {\bf u }_m + A_{sc} \left[ \tilde{{\bf u }}_1 e^{\sigma_{sc} t + i \omega_{sc} t} +   \tilde{\bf u }_{-1} e^{\sigma_{sc} t  -i \omega_{sc} t} \right],
\ee  
where ${\bf u }_m$ is the {\em mean flow} defined by phase-averaging over the cycle, $\tilde{{\bf u }}_1$ is a pseudo-eigenvector which is normalized by the condition  $||\tilde{{\bf u }}_1|| = 1/\sqrt{2}$, 
$A_{sc}$ is an amplitude parameter directly related to the energy of the oscillating flow, and $\lambda_{sc} = \sigma_{sc} + i \omega_{sc}$ is a pseudo-eigenvalue which depends upon the parameter $A_{sc}$. 
Introducing the decomposition (\ref{SC}) into the Navier-Stokes equations we are left 
with the following problem:

@@

Injecting and developing up to linear order leads to the following 
\begin{equation}
\begin{array}{ll}
\displaystyle
& {\cal NS}(  {\bf u }_0^g ) - [ {\cal C}( {\bf u }_{1,r}^g,{\bf u }_{1,r}^g) +  {\cal C}( {\bf u }_{1,i}^g,{\bf u }_{1,i}^g) ]
\\
+& {\cal L}_{{\bf u }_0^g}(\delta {\bf u }_0) - 2 [ {\cal C}( {\bf u }_{1,r}^g,\delta {\bf u }_{1,r}) +  {\cal C}( {\bf u }_{1,i}^g,\delta {\bf u }_{1,i}) ] = 0;
\\
\\
&  {\cal L}_{{\bf u }_0^g}( {\bf u }_{1,r}^g) - \omega^g {\bf u }_{1,i}^g 
 \\
  +& {\cal L}_{{\bf u }_0^g}( \delta {\bf u }_{1,r}^g) 
   -  {\cal C}( \delta {\bf u }_{0}^g, {\bf u }_{1,r})
   - \delta \omega^g \delta {\bf u }_{1,i}^g 
 - \omega^g \delta {\bf u }_{1,r} = 0 ;
 \\
\\
&  {\cal L}_{{\bf u }_0^g}({\bf u }_{1,i}^g) + \omega^g {\bf u }_{1,i}^g
 \\
  +& {\cal L}_{{\bf u }_0^g}( \delta {\bf u }_{1,i}^g) 
  -  {\cal C}( \delta {\bf u }_{0}^g, {\bf u }_{1,i}) 
  + \delta \omega^g \delta {\bf u }_{1,i}^g 
+ \omega^g \delta {\bf u }_{1,r} = 0 ;
\\
\\
& F_y({\bf u }_{1,i}^g) + F_y(\delta {\bf u }_{1,i}) = 0.
 \end{array}
\end{equation}

@@

\section*{Appendix A : details on mesh convergence}
\begin{figure}
\includegraphics[width=.9 \linewidth]{Cylinder_Sigma_Re_TESTMESHES.png}
\includegraphics[width=.9 \linewidth]{Cylinder_Strouhal_Re_TESTMESHES.png}
\caption{Growth rate $\sigma$ $(a)$  and Strouhal number $St = (a \omega)/(2\pi U_0)$ ($b$) as function of Reynolds number FOR 8 DIFFERENT MESHES}
\label{fig:MeshConvergence}
\end{figure}

\begin{figure*}
\small
\begin{lstlisting}
> baseflow=SF_Init('Mesh_Cylinder_Large.edp');
      ### INITIAL MESH CREATED WITH np = 2207 points
> baseflow=SF_BaseFlow(baseflow,'Re',1);
      ### FUNCTION SF_BaseFlow : computing base flow for Re = 1
      # Base flow converged in 6 iterations ; Drag = 5.9876; Lx = 0.501
> baseflow=SF_BaseFlow(baseflow,'Re',10);
      ### FUNCTION SF_BaseFlow : computing base flow for Re = 10
      # Base flow f converged in 5 iterations ; Drag = 1.4519; Lx = 0.73191
> baseflow=SF_BaseFlow(baseflow,'Re',60);
      ### FUNCTION SF_BaseFlow : computing base flow for Re = 60
      # Base flow converged in 6 iterations ; Drag = 0.66348; Lx = 3.6989

> baseflow=SF_Adapt(baseflow,'Hmax',10,'InterpError',0.01);
     ### ADAPT mesh to base flow ; InterpError = 0.01  ; Hmax = 10
      #   Number of points np = 1352 ; Ndof = 11862
      #  h_min, h_max : 0.018824 , 10.2895
      #  h_(A,B,C,D) : 0.027467 , 0.28644 , 0.45948 , 0.89245
      ### FUNCTION SF_BaseFlow : computing base flow for Re = 60
      # Base flow converged in 5 iterations ; Drag = 0.65805; Lx = 4.0737

> plotFF(baseflow,'mesh');
\end{lstlisting}
\normalsize
\caption{Illustration of the procedure to compute a base flow and an adapted mesh using StabFem (from script {\em SCRIPT\_CYLINDER.m}. }
\label{ill_bsflow}
\end{figure*}

\begin{figure*}
\small
\begin{lstlisting}
> [ev,em] = SF_Stability(baseflow,'shift',0.04+0.76i,'nev',1,'type','D');
> [baseflow,em]=SF_Adapt(baseflow,em,'Hmax',10,'InterpError',0.005);  
      ### ADAPT mesh to base flow AND MODE   ; InterpError = 0.005  ; Hmax = 10
      #   Number of points np = 13458 ; Ndof = 120030
      #  deltamin, deltapax : 0.0089891 , 13.367
      #  delta_(A,B,C,D) : 0.020536 , 0.082852 , 0.07316 , 0.11914
      ### FUNCTION SF_BaseFlow : computing base flow for Re = 60
      # Base flow converged in 4 iterations ; Drag = 0.64695; Lx = 4.0715
      ### FUNCTION SF_Stability : computation of 1 eigenvalues/modes (DIRECT) with FF solver
      # Stability calculation converged in 4 iterations, lambda = 0.047152+0.74808i.

> [ev,em] = SF_Stability(baseflow,'shift',0.04+0.76i,'nev',1,'type','S');
> [baseflow,em]=SF_Adapt(baseflow,em,'Hmax',10,'InterpError',0.005);
      ### ADAPT mesh to base flow AND SENSITIVITY ; InterpError = 0.005  ; Hmax = 10
      #   Number of points np = 4355 ; Ndof = 38559
      #  deltamin, deltapax : 0.0058536 , 13.1164
      #  delta_(A,B,C,D) : 0.01049 , 0.091861 , 0.13319 , 0.89372
      ### FUNCTION SF_BaseFlow : computing base flow for Re = 60
      # Base flow converged in 3 iterations ; Drag = 0.64758; Lx = 4.0725
      ### FUNCTION SF_Stability : computation of 1 eigenvalues/modes (DIRECT) with FF solver
      # Stability calculation converged in 3 iterations (D+A+S), lambda = 0.047165+0.74823i.

\end{lstlisting}
\normalsize

\caption{Illustration of the procedure to adapt the mesh for stability calculations with StabFem (from script {\em SCRIPT\_CYLINDER.m}. }
\label{stabmesh}
\end{figure*}

>>>>

%\begin{figure}
%\begin{center}
%\includegraphics[width=.9 \linewidth]{Cylinder_SC100_EnergySigma.png}
%\includegraphics[width=.9 \linewidth]{Cylinder_SC100_CySigma.png}
%\end{center}
%\caption{Self-consistent approach for the wake of a cyrcular cilinder: results at $Re =  100$.}
%\label{fig:SC100}
%\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=.9 \linewidth]{toto.png}
\end{center}
\caption{Self-consistent approach for the wake of a circular cilinder:  description of the MEAN FLOW for  $Re = 
100$.}
\label{fig:MF60}
\end{figure}

We now briefly review the self-consistent approach as introduced by Manti\v{c}-Lugo et 
al \cite{MLugo2014}. The authors adopted a pseudo-eigenmode expansion of the flow 
as follows: 
\be{SC}
{\bf u } = {\bf u }_m + A_{sc} \left[ \tilde{{\bf u }}_1 e^{\sigma_{sc} t + i \omega_{sc} t} +   \tilde{\bf u }_{-1} e^{\sigma_{sc} t  -i \omega_{sc} t} \right],
\ee  
where ${\bf u }_m$ is the {\em mean flow} defined by phase-averaging over the cycle, $\tilde{{\bf u }}_1$ is a pseudo-eigenvector which is normalized by the condition  $||\tilde{{\bf u }}_1|| = 1/\sqrt{2}$, 
$A_{sc}$ is an amplitude parameter directly related to the energy of the oscillating flow, and $\lambda_{sc} = \sigma_{sc} + i \omega_{sc}$ is a pseudo-eigenvalue which depends upon the parameter $A_{sc}$. 
Introducing the decomposition (\ref{SC}) into the Navier-Stokes equations we are left 
with the following problem:

@@

Injecting and developing up to linear order leads to the following

TO BE CORRECTED !

\begin{equation}
\begin{array}{ll}
\displaystyle
& {\cal NS}(  {\bf u }_0^g ) - [ {\cal C}( {\bf u }_{1,r}^g,{\bf u }_{1,r}^g) +  {\cal C}( {\bf u }_{1,i}^g,{\bf u }_{1,i}^g) ]
\\
+& {\cal L}_{{\bf u }_0^g}(\delta {\bf u }_0) - 2 [ {\cal C}( {\bf u }_{1,r}^g,\delta {\bf u }_{1,r}) +  {\cal C}( {\bf u }_{1,i}^g,\delta {\bf u }_{1,i}) ] = 0;
\\
\\
&  {\cal L}_{{\bf u }_0^g}( {\bf u }_{1,r}^g) - \omega^g {\bf u }_{1,i}^g 
 \\
  +& {\cal L}_{{\bf u }_0^g}( \delta {\bf u }_{1,r}^g) 
   -  {\cal C}( \delta {\bf u }_{0}^g, {\bf u }_{1,r})
   - \delta \omega^g \delta {\bf u }_{1,i}^g 
 - \omega^g \delta {\bf u }_{1,r} = 0 ;
 \\
\\
&  {\cal L}_{{\bf u }_0^g}({\bf u }_{1,i}^g) + \omega^g {\bf u }_{1,i}^g
 \\
  +& {\cal L}_{{\bf u }_0^g}( \delta {\bf u }_{1,i}^g) 
  -  {\cal C}( \delta {\bf u }_{0}^g, {\bf u }_{1,i}) 
  + \delta \omega^g \delta {\bf u }_{1,i}^g 
+ \omega^g \delta {\bf u }_{1,r} = 0 ;
\\
\\
& F_y({\bf u }_{1,i}^g) + F_y(\delta {\bf u }_{1,i}) = 0.
 \end{array}
\end{equation}

@@

\section*{Appendix A : details on mesh convergence}

TO BE FINISHED
